### 2023-08-02 README.Rmd
230802_changes

The goal of validateHOT is to validate the results of your validation
task (also known as holdout task). A validation task is essential to
make sure that your collected data of a MaxDiff, CBC, or ACBC are valid
and can also predict outside task that were not included in estimating
your utility scores. Although commercial studies often do not include a
validation/holdout task (Yang et al., 2018), it is highly recommended to do so
(Orme, 2015; Rao, 2014). This validation/ holdout task does not only
help to check whether everything went right during data collection but
also to determine your final model. <code>validatHOT</code> provides
some of the relevant metrics to test the performance of your data in
predicting a holdout task. In [Sawtooth
Software's](https://sawtoothsoftware.com/) CBC a fixed validation/
holdout task is automatically implemented. If you conduct a MaxDiff or
ACBC study these have to be programmed by yourself.

@@

For all 3 key areas, the <code>createHOT</code> function is essential.
This function creates the total utilities for each alternative in the
validation/holdout task and also in the simulation method, respectively.
Rao (2014, p. 82) mentions the additive utility model stating that the
total utility of a profile or conjoint model is the sum of its attribute
levels. <code>createHOT</code> will do exactly this for you.

@@

<li><code>hitrate</code>: creates the Hit Rate (correctly predicted
choices) of your validation task. The output will contain the chance
level in your validation task ($\frac{1}{alternatives}$) in percentage.
The number of correctly predicted participants' choices as well as the
percentage of how many choices were correctly predicted. If you specify
an optional <code>Group</code> argument the output is split by
groups.</li>

<li><code>kl</code>: Kullback-Leibler-Divergence which measures the
divergence between the actual choice distribution and the predicted
choice distribution (Ding et al., 2011; Drost, 2018). Output provides both
divergence between predicted from observed and observed from predicted
due to the asymmetry of the Kullback-Leibler divergence. If you specify
an optional <code>Group</code> argument the output is split by
groups.</li>

<li><code>mae</code>: average absolute error, i.e., deviation between
predicted and stated choice share. If you specify an optional
<code>Group</code> argument the output is split by groups.</li>

<li><code>medae</code>: since the averaged absolute error can be highly
influenced by the If you specify an optional <code>Group</code> argument
the output is split by groups.</li>

<li><code>mhp</code>: averaged hit probability of participant's actual
choice in the validation/ holdout task. If you specify an optional
<code>Group</code> argument the output is split by groups.</li>

<li><code>rmse</code>: provides the root-mean-squared error of deviation
between predicted and stated choice share. If you specify an optional
<code>Group</code> argument the output is split by groups.</li>

@@

<li><code>accuracy</code>: defined as
$\frac{TP + TN}{TP + FP + TN + FN}$. If you specify an optional
<code>Group</code> argument the output is split by groups.</li>

<li><code>f1</code>: defined as
$\frac{2 * precision * recall}{precision + recall}$ or stated
differently by Burger (2018) $\frac{2TP}{2TP + FP + FN}$. If you
specify an optional <code>Group</code> argument the output is split by
groups.</li>

<li><code>precision</code>: defined as $\frac{TP}{TP + FP}$. If you
specify an optional <code>Group</code> argument the output is split by
groups.</li>

<li><code>recall</code>: defined as $\frac{TP}{TP + FN}$. If you specify
an optional <code>Group</code> argument the output is split by
groups.</li>

<li><code>specificity</code>: defined as $\frac{TN}{TN + FP}$. If you
specify an optional <code>Group</code> argument the output is split by
groups.</li>

@@

<li><code>freqassort</code>: Inspired by the former <a href="https://github.com/cran/turfR" target="_blank">turfR</a> package, <code>freqassort</code>
will give you the averaged frequency, how many products the participants
will choose from your in the function determined potential assortment.
For the <code>method</code> argument you can decide between <code>method
= "threshold"</code> (if utility of product is larger than the utility
of <code>None</code>, it is marked as potential purchase option) and
<code>method = "First Choice"</code> (only product with highest utility
is considered. If its utility is above the utility of <code>None</code>,
it is marked as potential purchase option). If you specify an optional
<code>Group</code> argument the output is split by groups.</li>

<li><code>reach</code>: Inspired by the former
<a href="https://github.com/cran/turfR" target="_blank">turfR</a> package, <code>reach</code> will
give you the averaged percentage of how many participants you can reach
(at least one of the products resemble a purchase option) with your in the function determined
potential assortment. or the <code>method</code> argument you can decide
between <code>method = "threshold"</code> (if utility of product is
larger than the utility of <code>None</code>, it is marked as potential
purchase option) and <code>method = "First Choice"</code> (only product
with highest utility is considered. If its utility is above the utility
of <code>None</code>, it is marked as potential purchase option). If you
specify an optional <code>Group</code> argument the output is split by
groups.</li>

<li><code>shareofpref</code>: provides you the aggregated share of
preference, including the lower and upper confidence interval, which is
calculated according to the $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$. If
you specify an optional <code>Group</code> argument the output is split
by groups and provided in a list element.</li>

@@

We are teaching a preference measurement seminar for students. Often
these students did not have experience with *R* before or only sparsely.
We teach them to also validate their results and wanted to give them an
easy way on how to do this in *R*. Of course, there are other great
packages which are even faster in running (i.e., Hamner & Frasco, 2018), however,
these packages need some more data wrangling in order to use the
appropriate functions, which might be a burden or barrier for the one or
the other.

@@

As mentioned above, all the attributes are part-worth coded, therefore, we set <code>coding = c(0, 0, 0)</code>. Finally, we specify the method of prefernce measurement technique, which is <code>method = "CBC"</code> in our case and the column index of the actual participant's choice (<code>choice</code>). If you run the code, a data frame called <code>HOT</code> (short form of **H**old**o**ut **t**ask) will be returned to the global environment.

> &#10071; <code>validateHOT</code> is currently just taking indexes instead of column names, please be aware of this. However, you can easily find out the index by using the <code>names()</code> function or by using <code>which()</code> and <code>colnames()</code>, both functions are provided by the base package (R Core Team, 2023). For example, if you want to find out the column index of <code>id</code>, we could also use <code>which(colnames(CBC) == "ID")</code>.

@@

In the next step, we would like to see how well our model (from which we took the raw utilities) predict the choices in the validation/ holdout task. First, we will test the <code>hitrate()</code> function. We specify the <code>data</code>, the column index of <code>id</code>, the column index of the alternatives (<code>opts</code>; remember we have three alternatives + the *no-buy* alternative), and finally the column index of the choice. We can see that the hitrate of our example is `r round(hitrate(data = HOT, id = 1, opts = c(2:5), choice = 6)[3, 1], digits  = 3)`.

```{r}
hitrate(data = HOT, id = 1, opts = c(2:5), choice = 6)
```

@@

```{r}
mae(data = HOT, id = 1, opts = c(2:5), choice = 6)
```
To cover also one example on how to use the metrics of the confusion matrix, we could test whether our model overestimates the purchase behavior (our model predicts a *buy* although participant opts for a *no* buy) or underestimates it (i.e., model predicts *no-buy* but participant opts for a *buy*). We will test the accuracy of the model by running the <code>accuracy()</code> function.

```{r}
accuracy(data = HOT, id = 1, opts = c(2:5), None = 5, choice = 6)
```
Finally, let us test, how many participants would at least buy one of the three products, assuming that this is one potential assortment we would like to offer to our consumers. We will use the <code>reach()</code> function and use the threshold approach. To specify the bundles we are offering we use the <code>bundles</code> argument in our function.

```{r}
reach(data = HOT, id = 1, bundles = c(2:4), None = 5, method = "threshold")
```

@@

The next steps are the same as above. However, let us take a look at some examples in which we display the results per group. Let us again begin with the <code>hitrate()</code> function. To do so, we specify the column index of the grouping variable in the <code>Group</code> argument.

```{r}
hitrate(data = CBC, id = 1, opts = c(2:5), Group = 6, choice = 7)
```

In this case, the Grouping variable is just an integer. However, the output is different if it is a factor or if it is labelled data. To proof this we just quickly change <code>Group</code> in a factor by using the <code>factor()</code> function provided by R Core Team (2023).

@@

```{r}
mhp(data = CBC, id = 1, opts = c(2:5), Group = 6, choice = 7)
```

>>>>

The goal of validateHOT is to validate the results of your validation
task (also known as holdout task). A validation task is essential to
make sure that your collected data of a MaxDiff, CBC, or ACBC are valid
and can also predict outside task that were not included in estimating
your utility scores. Although commercial studies often do not include a
validation/ holdout task (Yang et al., 2018), it is highly recommended to do so
(Orme, 2015; Rao, 2014). This validation/ holdout task does not only
help to check whether everything went right during data collection but
also to determine your final model. <code>validatHOT</code> provides
some of the relevant metrics to test the performance of your data in
predicting a holdout task. In [Sawtooth
Software's](https://sawtoothsoftware.com/) CBC a fixed validation/
holdout task is automatically implemented. If you conduct a MaxDiff or
ACBC study these have to be programmed by yourself.

@@

For all 3 key areas, the <code>createHOT</code> function is essential.
This function creates the total utilities for each alternative in the
validation/ holdout task and also in the simulation method, respectively.
Rao (2014, p. 82) mentions the additive utility model stating that the
total utility of a profile or conjoint model is the sum of its attribute
levels. <code>createHOT</code> will do exactly this for you.

@@

<li><code>hitrate</code>: creates the Hit Rate (correctly predicted
choices) of your validation task. The output will contain the chance
level in your validation task ($\frac{1}{alternatives}$) in percentage.
The number of correctly predicted participants' choices as well as the
percentage of how many choices were correctly predicted. If you specify
an optional <code>group</code> argument the output is split by
groups.</li>

<li><code>kl</code>: Kullback-Leibler-Divergence which measures the
divergence between the actual choice distribution and the predicted
choice distribution (Ding et al., 2011; Drost, 2018). Output provides both
divergence between predicted from observed and observed from predicted
due to the asymmetry of the Kullback-Leibler divergence. If you specify
an optional <code>group</code> argument the output is split by
groups. Currently, you can choose between \eqn{log{_2}} and \eqn{log} as logarithm base.</li>

<li><code>mae</code>: average absolute error, i.e., deviation between
predicted and stated choice share. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

<li><code>medae</code>: since the averaged absolute error can be highly
influenced by the If you specify an optional <code>group</code> argument
the output is split by groups.</li>

<li><code>mhp</code>: averaged hit probability of participant's actual
choice in the validation/ holdout task. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

<li><code>rmse</code>: provides the root-mean-squared error of deviation
between predicted and stated choice share. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

@@

<li><code>accuracy</code>: defined as
$\frac{TP + TN}{TP + FP + TN + FN}$. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

<li><code>f1</code>: defined as
$\frac{2 * precision * recall}{precision + recall}$ or stated
differently by Burger (2018) $\frac{2TP}{2TP + FP + FN}$. If you
specify an optional <code>group</code> argument the output is split by
groups.</li>

<li><code>precision</code>: defined as $\frac{TP}{TP + FP}$. If you
specify an optional <code>group</code> argument the output is split by
groups.</li>

<li><code>recall</code>: defined as $\frac{TP}{TP + FN}$. If you specify
an optional <code>group</code> argument the output is split by
groups.</li>

<li><code>specificity</code>: defined as $\frac{TN}{TN + FP}$. If you
specify an optional <code>group</code> argument the output is split by
groups.</li>

@@

<li><code>freqassort</code>: Inspired by the former <a href="https://github.com/cran/turfR" target="_blank">turfR</a> package, <code>freqassort</code>
will give you the averaged frequency, how many products the participants
will choose from your in the function determined potential assortment.
Again, you have to define a `none` alternative. `freqassort` uses the *threshold* approach,
meaning if the utility of one product is above the utility of `none`,
it is marked as potential purchase option. If you specify an optional
<code>group</code> argument the output is split by groups.</li>

<li><code>reach</code>: Inspired by the former
<a href="https://github.com/cran/turfR" target="_blank">turfR</a> package, <code>reach</code> will
give you the averaged percentage of how many participants you can reach
(at least one of the products resemble a purchase option) with your in the function determined
potential assortment. `reach` also uses the *threshold* approach (see above).</li>

<li><code>shareofpref</code>: provides you the aggregated share of
preference, including the lower and upper confidence interval, which is
calculated according to the $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$. If
you specify an optional <code>group</code> argument the output is split
by groups and provided in a list element.</li>

@@

We are teaching a preference measurement seminar for students. Often
these students did not have experience with *R* before or only sparsely.
We teach them to also validate their results and wanted to give them an
easy way on how to do this in *R*. Of course, there are other great
packages which are faster in running (i.e., Hamner & Frasco, 2018), however,
these packages need some more data wrangling in order to use the
appropriate functions, which might be a burden or barrier for the one or
the other.

@@

As mentioned above, all the attributes are part-worth coded, therefore, we set <code>coding = c(0, 0, 0)</code>. Finally, we specify the method of preference measurement technique, which is <code>method = "CBC"</code> in our case and the column index of the actual participant's choice (<code>choice</code>). If you run the code, a data frame called <code>HOT</code> (short form of **H**old**o**ut **t**ask) will be returned to the global environment.

> &#10071; <code>createHOT</code> is currently just taking indexes instead of column names, please be aware of this. However, you can easily find out the index by using the <code>names()</code> function or by using <code>which()</code> and <code>colnames()</code>, both functions are provided by the base package (R Core Team, 2023). For example, if you want to find out the column index of <code>id</code>, we could also use <code>which(colnames(CBC) == "ID")</code>.

@@

In the next step, we would like to see how well our model (from which we took the raw utilities) predict the choices in the validation/ holdout task. First, we will test the <code>hitrate()</code> function. We specify the <code>data</code>, the column names of the alternatives (<code>opts</code>; remember we have three alternatives + the *no-buy* alternative), and finally the choice. We can see that the hitrate of our example is `r round(as.numeric(hitrate(data = HOT, opts = c(2:5), choice = choice)[3, 2]), digits  = 3)`.

```{r}
hitrate(data = HOT, opts = c(2:5), choice = choice)
```

@@

```{r}
mae(data = HOT, opts = c(2:5), choice = choice)
```
To cover also one example on how to use the metrics of the confusion matrix, we could test whether our model overestimates the purchase behavior (our model predicts a *buy* although participant opts for a *no* buy) or underestimates it (i.e., model predicts *no-buy* but participant opts for a *buy*). We will test the accuracy of the model by running the <code>accuracy()</code> function.

```{r}
accuracy(data = HOT, opts = c(2:5), choice = choice, none = None)
```
Finally, let us test, how many participants would at least buy one of the three products, assuming that this is one potential assortment we would like to offer to our consumers. We will use the <code>reach()</code> function and use the threshold approach. To specify the bundles we are offering we use the <code>opts</code> argument in our function.

```{r}
reach(data = HOT, opts = c(Option_1:Option_3), none = None)
```

@@

The next steps are the same as above. However, let us take a look at some examples in which we display the results per group. Let us again begin with the <code>hitrate()</code> function. To do so, we specify the column name of the grouping variable in the <code>group</code> argument.

```{r}
hitrate(data = CBC, opts = c(Option_1:None), choice = choice, group = Group)
```

In this case, the Grouping variable is just an integer. However, the output is different if it is a factor or if it is labelled data. To proof this we just quickly change <code>group</code> in a factor by using the <code>factor()</code> function provided by R Core Team (2023).

@@

```{r}
mhp(data = CBC, opts = c(Option_1:None), choice = choice, group = Group)
```
